### 2017-06-21 manuscript.tex
add some refs and begin methods

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Quantitative comparison of the potential for viral escape from broad and narrow anti-influenza antibodies} % Please use "title case" (capitalize all terms in the title except conjunctions, prepositions, and articles).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Michael B. Doud\textsuperscript{1,2,3},
Jesse D. Bloom\textsuperscript{1,2*},
\\
\bigskip
\textbf{1} Basic Sciences and Computational Biology Program, Fred Hutchinson Cancer Research Center, Seattle, WA, USA
\\
\textbf{2} Department of Genome Sciences, University of Washington, Seattle, WA,  USA
\\
\textbf{3} Medical Scientist Training Program, University of Washington, Seattle, WA, USA
\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
There is great interest in broad antibody-based immunity against influenza virus.
Even for antibodies recognizing broad swaths of influenza strains, viral genetic variation can influence neutralization sensitivity.
To better understand the potential for viral resistance to antibodies, there is need to evaluate not just whether mutations can influence sensitivity to an antibody, but also to quantify how many mutations do this and by how much, and to make these comparisons across antibodies.
Here we comprehensively map mutations that increase viral resistance to the broadly neutralizing anti-influenza antibody FI6v3, and compare the results to those previously obtained for narrow strain-specific antibodies.
We find that although there are some mutations that enhance resistance to FI6v3, these mutations are both rare and modest in their effect sizes.
We formalize these observations into a framework for quantitative comparison of viral escape potential between different antibodies.
Overall, our work demonstrates that a virus's absolute capacity to escape an antibody can be quantified, and that influenza is in fact extremely limited in its ability to escape some antibodies.

@@

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
this is the introduction

@@

\subsubsection*{methods subsubsection 1}

\subsubsection*{Data availability and source code}
Deep sequencing data has been deposited at the Sequence Read Archive under BioSample accession ().

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Quantitative comparison of the potential for viral escape from broad and narrow anti-influenza antibodies} % Please use "title case" (capitalize all terms in the title except conjunctions, prepositions, and articles).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Michael B. Doud\textsuperscript{1,2,3},
Juhye M. Lee\textsuperscript{1,2,3},
Jesse D. Bloom\textsuperscript{1,2*},
\\
\bigskip
\textbf{1} Basic Sciences and Computational Biology Program, Fred Hutchinson Cancer Research Center, Seattle, WA, USA
\\
\textbf{2} Department of Genome Sciences, University of Washington, Seattle, WA,  USA
\\
\textbf{3} Medical Scientist Training Program, University of Washington, Seattle, WA, USA
\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
There is great interest in broad antibody-based immunity against influenza virus.
Even for antibodies recognizing broad swaths of influenza strains, viral genetic variation can influence neutralization sensitivity.
To better understand the potential for viral resistance to antibodies, there is a need to evaluate not just whether mutations can influence sensitivity to an antibody, but rather to comprehensively quantify how many mutations do this and by how much, and to make these comparisons across antibodies.
Here we comprehensively map mutations that increase viral resistance to the broadly neutralizing anti-influenza antibodies FI6v3, C179, and S139/1, and compare the results to those previously obtained for narrow strain-specific antibodies.
We find that although there are some mutations that enhance resistance to the stalk-binding antibody FI6v3 \comment{and C179?}, these mutations are both rare and modest in their effect sizes.
\comment{..?... In contrast, viral escape from S139/1 ... looks more like the narrow head binders than the broad stalk binders?}
We formalize these observations into a framework for quantitative comparison of viral escape potential between different antibodies.
Overall, our work demonstrates that a virus's absolute capacity to escape an antibody can be quantified, and that influenza is in fact extremely limited in its ability to escape some antibodies \comment{targeting the stalk... and also extremely adept at escaping another broadly-reactive antibody targeting the head... }.

@@

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}

Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.

% blurbs about the specific antibodies we use here:
C179 isolation and escape mutant selection \cite{okuno1993common}.
C179 structure with 1957 H2, binding data to many strains, and citations to older studies demonstrating C179 cross-neutralization to H1, H2, H5, H6, and H9:  \cite{dreyfus2013structure}.
FI6v3 was isolated... \cite{corti2011neutralizing}.
S139/1 isolation and selection of escape mutants from Aichi H3, Adachi H2, and WSN H1: \cite{yoshida2009cross}.
S139/1 structure with Victoria75 H3 and binding/neutralization data: \cite{lee2012heterosubtypic}.

Recently we have described new technologies... \cite{doud2017complete, dingens2017comprehensive}.

@@

\subsubsection*{Antibodies}
FI6v3 was expressed and purified by the Fred Hutchinson Cancer Research Center protein expression core \comment{i think?}.
C179 was purchased from Takara Bio Inc (Catalog \# M145).
S139/1 heavy and light chain variable sequences were obtained from PDB ID 4GMS~\cite{lee2012heterosubtypic} and expressed and purified by the Fred Hutchinson Cancer Research Center protein expression core.

\subsubsection*{Mutant virus selections with antibody}
Library selections, deep sequencing, and computation of mutation differential selection were performed as previously described~\cite{doud2017complete}. Briefly, ...

\subsubsection*{Computation of the fraction $\phi_{r,a}$ of each mutation that escapes antibody neutralization}
\comment{depending on how much of paper focus is on this new metric, some of this rationale might be better placed in the results section}

Mutation differential selection values $s_{r,a}$ reflect the enrichment of a mutation in an antibody-selected sample relative to a mock-selected control. 
The extent of these mutation enrichments are dependent on the stringency of the neutralization: as the concentration of an antibody used to neutralize the mutant virus library is increased, a larger proportion of the virus library is neutralized, and escape mutants become more enriched in the antibody-selected sample relative to the control sample~\cite{doud2017complete}.
Consequently, due to differences in the concentrations and potencies of various antibodies used in mutational antigenic profiling, quantitative comparisons of the effects of mutations on neutralization escape between various antibodies is confounded by differences in neutralization stringency across experiments.

We defined a new metric that accounts for the neutralization stringency of each experiment so that quantitative comparisons can be made between experiments with varying neutralization stringencies.
This new measure, $\phi_{r,a}$, is roughly analogous to the fraction of mutant viruses carrying mutation $a$ at site $r$ that escape neutralization by a given antibody.
The computation of $\phi_{r,a}$ incorporates information about the stringency of neutralization. 
We define this as  $\gamma$ = (100 - \% neutralization)/100), the fraction of library remaining infectious after antibody neutralization measured by qRT-PCR for each antibody-selection experiment as previously described~\cite{doud2017complete}.

Define $\hat{n}_{r,a}^{sel}$ and $\hat{n}_{r,a}^{mock}$ as the number of error-corrected counts of amino-acid mutant $a$ at site $r$ as defined in~\cite{doud2017complete}, and $N_r^{sel}$ and $N_r^{mock}$ be the total counts at site $r$, as specified for either an antibody selection or matched mock condition.
Let $P$ be a pseudocount (default to 5 in these analyses) added to amino-acid counts, and let $f_{r}^{sel}$ and $f_{r}^{mock}$ be relative depths of the selected and mock samples at site $r$ for the purposes of scaling the pseudocount by relative read depth as previously described~\cite{doud2017complete}.

Define the pseudocount-adjusted frequency $\rho$ of mutant $a$ at site $r$ in selected and mock samples to be:
$$\rho_{r,a}^{sel} = \frac{n_{r,a}^{sel}+f_r^{sel}\times P}{N_r^{sel}+f_r^{sel}\times  P}$$
$$\rho_{r,a}^{mock} = \frac{n_{r,a}^{mock}+f_r^{mock}\times P}{N_r^{mock}+f_r^{mock}\times  P}$$

We define $\phi_{r,a}$ as the net `fraction' of mutant $a$ at site $r$ escaping neutralization (above the average escape fraction $\gamma$), and compute this as:
$$\phi_{r,a} = \frac{\gamma \times \rho_{r,a}^{sel}}{\rho_{r,a}^{mock}} - \gamma$$

By this definition, mutations with no effect on antibody escape will have $\phi_{r,a} =0$, and mutations completely escaping antibody will have $\phi_{r,a} \approx 1$.
Care should be taken in interpreting the absolute values of $\phi$, however, since the choice of pseudocount affects the range of $\phi$ values (but not, importantly, the relationships of $\phi$ distributions across antibodies).
We have found that a value of $P=5$ pseudocounts results in $\phi$ values that are roughly bounded from 0 to 1, and consistently use $P=5$ all analyses.

\subsubsection*{Data availability and source code}
Deep sequencing data has been deposited at the Sequence Read Archive under BioSample accession (\comment{should probably deposit all the new data into the existing biosample SAMN05789126 (Sample name: WSN/1933 HA mutant libraries selected by monoclonal antibodies), which is under bioproject PRJNA309339}).
